---
title: "Assignment5_Part1"
output:
  ioslides_presentation: default
  beamer_presentation: default
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

## Team members
- Yue Wu 
- YinChia Huang 
- Francesco Ignazio Re

## Outline
- Introduction
- Data cleaning
- Data visualization
- Statistic Analysis

## Introduction
The purpose of this project is for us to exercise, extend, and demonstrate our abilities to use the ideas and tools of the Tidyverse. We are assigned data from NOAA Weather Station buoy 46035 at 57.026 N 177.738 W in the NOAA National
Data Buoy Center. Our goal is to obtain data, clean, organize, and explore it. 

## Data cleaning
We only focused on Air Temperature and Sea Temperature from 1988 to 2017. However, data of 2012 and 2013 are missing, which we replaced by the data of  nearby buoy 46070, assuming their temperatures should be pretty close.
In addition to 2012 and 2013, there are a few missing values in other years. Every missing data has value 99.0 or 999.0. As a result, we deleted them, otherwise they would be regarded as outliers.  


```{r include=FALSE}
library(tidyverse)
library(stringr)
library("reshape2")
library("ggplot2")


url1 <- "http://www.ndbc.noaa.gov/view_text_file.php?filename=46035h"
url2 <- "http://www.ndbc.noaa.gov/view_text_file.php?filename=46070h"
url3 <- ".txt.gz&dir=data/historical/stdmet/"
years <- c(1988:2017)
urls1 <- str_c(url1, years, url3, sep = "")
urls2 <- str_c(url2, years, url3, sep = "")
filenames <- str_c("mr", years, sep = "")
N <- length(years)

#create files for each year 
i<-1
while (i <= N){
  
  if(i==25|i==26){
    suppressMessages(
      assign(filenames[i], read_table(urls2[i], col_names = TRUE)))
  }
  else{
    suppressMessages(
      assign(filenames[i], read_table(urls1[i], col_names = TRUE)))
  }
  file <- get(filenames[i])
  
  #add 19 in front of 2 digits year(from 1988 to 1998)
  colnames(file)[1] <-"YYYY"
  if(i>=1 && i<=11){
    file$YYYY<-paste("19", file$YYYY, sep = "")
  }
  #only select time, air temp and water temp
  file <- subset(file, select=c("YYYY", "MM","DD","hh","ATMP","WTMP"))
  file$ATMP<-as.numeric(file$ATMP)
  file$WTMP<-as.numeric(file$WTMP)
  file <- subset(file, ATMP!= 999.0 & ATMP!= 99.0)
  file <- subset(file, WTMP!= 999.0 & WTMP!= 99.0)
  annualMean_AT <- mean(file$ATMP, na.rm=TRUE)
  annualMean_WT <- mean(file$WTMP, na.rm=TRUE)
  
  if(i==1){
    MR <- file
    MR_DailyNoon <- subset(file, hh==12)
    annualMeans_AT <- annualMean_AT
    annualMeans_WT <- annualMean_WT
  }
  else{
    MR <- rbind.data.frame(MR, file)
    MR_DailyNoon <- rbind.data.frame(MR_DailyNoon, subset(file, hh==12))
    annualMeans_AT <- rbind.data.frame(annualMeans_AT, annualMean_AT)
    annualMeans_WT <- rbind.data.frame(annualMeans_WT, annualMean_WT)
  }
  i<-i+1
}
MR$FullTime <-strptime(with(MR, paste(YYYY, MM, DD, hh, sep="-")), format="%Y-%m-%d-%H")
MR_DailyNoon$FullTime <-strptime(with(MR_DailyNoon, paste(YYYY, MM, DD, hh, sep="-")), format="%Y-%m-%d-%H")

```


## Data visualization
```{r echo=FALSE}
ggplot(MR_DailyNoon, aes(x = FullTime)) + 
  geom_line(aes(y = ATMP, colour ="Air Temperature")) + 
  geom_line(aes(y = WTMP, colour = "Water Temperature")) +
  scale_colour_manual("", 
                      breaks = c("Air Temperature", "Water Temperature"),
                      values = c("pink", "blue"))+
  xlab("Time") +
  scale_y_continuous("Temperatura (C)", limits = c(0,10)) + 
  labs(title="A time series composed of 30 years of daily Air Temperature
and Sea Temperature readings recorded at noon")
  
```

## Has the mean temperature changed over the past 30 years?
```{r echo=FALSE}
#create annual temprature table, which containing means of temperatures in each year
annualTemprature <- data.frame(c(1988:2017), annualMeans_AT, annualMeans_WT)
colnames(annualTemprature) <- c("Year", "Air Temperature", "Water Temperature")

ggplot(annualTemprature, aes(x = Year)) + 
  geom_line(aes(y = annualMeans_AT , colour="Air Temperature")) + 
  geom_line(aes(y = annualMeans_WT , colour = "Water Temperature")) +
   scale_colour_manual("", 
                      breaks = c("Air Temperature", "Water Temperature"),
                      values = c("pink", "blue"))+
  ylab(label="Celsius degrees") + 
  xlab("Time")+
  labs(title="Time series composed from 1988 to 2017 recording annual mean of Air Temperature
and Sea Temperature")
```

## Statistical Test- Air temperature :1988_noon V.S 2017_noon 
```{r include=FALSE}
MR_1988_DailyNoon <- subset(MR_DailyNoon, YYYY==1988)[,c(2:6)]  
MR_2017_DailyNoon <- subset(MR_DailyNoon, YYYY==2017)[,c(2:6)]
colnames(MR_1988_DailyNoon) <- c("MM","DD","hh","ATMP1988","WTMP1988")
colnames(MR_2017_DailyNoon) <- c("MM","DD","hh","ATMP2017","WTMP2017")

#combine 1988 and 2017 into one table
MR_1988VS2017_DailyNoon <- as.data.frame(left_join(MR_1988_DailyNoon,MR_2017_DailyNoon))
```

```{r echo=FALSE}
#t test on air tmp
t.test(MR_1988VS2017_DailyNoon["ATMP1988"],MR_1988VS2017_DailyNoon["ATMP2017"])    
```
p=0 < 0.05, we can reject the null hypothesis that daily Air Temperature recorded at noonin in 1988 and 2017 have the same mean


## Statistical Test - Water temperature :1988_noon V.S 2017_noon
```{r}
#t test on water tmp
t.test(MR_1988VS2017_DailyNoon["WTMP1988"],MR_1988VS2017_DailyNoon["WTMP2017"]) 
```
p=0 < 0.05, we can reject the null hypothesis that daily Water Temperature recorded at noonin in 1988 and 2017 have the same mean

## Conclusion : There are significant changes in the past 30 years


## Since we only use one sample per day day out of 24 daily hourly temperature readings. Has your sampling affected your evaluation of temperature change? In what way? Explain and demonstrate

## Statistical Test - air temperature
Method 1:  Test the difference based on choice of time of the day
Ramdonly chose 20:00 to test
```{r include=FALSE}
MR_Daily20 <- subset(MR, hh==20)[,-c(4,7)]
colnames(MR_Daily20) <- c("YYYY","MM","DD","ATMP20","WTMP20")
colnames(MR_DailyNoon) <- c("YYYY","MM","DD","hh","ATMP12","WTMP12","Full Time")
MR_NoonVS20 <- as.data.frame(left_join(MR_DailyNoon[,-c(4,7)],MR_Daily20))
```


```{r echo=FALSE}
#test on air tmp
t.test(MR_NoonVS20["ATMP12"],MR_NoonVS20["ATMP20"])   
```
p=0.3262 > 0.05, we cannot reject the null hypothesis that daily noon and 20:00 air tmp have the same mean

## Statistical Test - water temperature

```{r echo=FALSE}
#test on water tmp
t.test(MR_NoonVS20["WTMP12"],MR_NoonVS20["WTMP20"])
```
p=0.7733 > 0.05, we cannot reject the null hypothesis that daily noon and 20:00 water tmp have the same mean

## Statistical Test 
Method 2:  Prove that choice of the hour won't effect the test result 
```{r}
#Demonstration:
MR_Daily00 <- subset(MR, hh=="00")[,-c(4,7)]
colnames(MR_Daily00) <- c("YYYY","MM","DD","ATMP00","WTMP00")
MR_Daily02 <- subset(MR, hh=="02")[,-c(4,7)]
colnames(MR_Daily02) <- c("YYYY","MM","DD","ATMP02","WTMP02")
MR_Daily04 <- subset(MR, hh=="04")[,-c(4,7)]
colnames(MR_Daily04) <- c("YYYY","MM","DD","ATMP04","WTMP04")
MR_Daily06 <- subset(MR, hh=="06")[,-c(4,7)]
colnames(MR_Daily06) <- c("YYYY","MM","DD","ATMP06","WTMP06")
MR_Daily08 <- subset(MR, hh=="08")[,-c(4,7)]
colnames(MR_Daily08) <- c("YYYY","MM","DD","ATMP08","WTMP08")
MR_Daily10 <- subset(MR, hh==10)[,-c(4,7)]
colnames(MR_Daily10) <- c("YYYY","MM","DD","ATMP10","WTMP10")
MR_Daily14 <- subset(MR, hh==14)[,-c(4,7)]
colnames(MR_Daily14) <- c("YYYY","MM","DD","ATMP14","WTMP14")
MR_Daily16 <- subset(MR, hh==16)[,-c(4,7)]
colnames(MR_Daily16) <- c("YYYY","MM","DD","ATMP16","WTMP16")
MR_Daily18 <- subset(MR, hh==18)[,-c(4,7)]
colnames(MR_Daily18) <- c("YYYY","MM","DD","ATMP18","WTMP18")
MR_Daily22 <- subset(MR, hh==22)[,-c(4,7)]
colnames(MR_Daily22) <- c("YYYY","MM","DD","ATMP22","WTMP22")
MR_every2hrs <- left_join(MR_Daily00,MR_Daily02) %>%
                left_join(.,MR_Daily04) %>%
                left_join(.,MR_Daily06) %>%
                left_join(.,MR_Daily08) %>%
                left_join(.,MR_Daily10) %>%
                left_join(.,MR_DailyNoon[,-c(4,7)]) %>%
                left_join(.,MR_Daily14) %>%
                left_join(.,MR_Daily16) %>%
                left_join(.,MR_Daily18) %>%
                left_join(.,MR_Daily20) %>%
                left_join(.,MR_Daily22)
MR_every2hrs <- as.data.frame(MR_every2hrs)

```

```{r}
summary(MR_every2hrs[,-c(1:3)])
```

## Statistical Test- air temperature
Method 3. Test the difference if we use all daily data, instead of specific hours per day
```{r include=FALSE}
MR_1988 <- subset(MR, YYYY==1988)[,c(2:6)]  
MR_2017 <- subset(MR, YYYY==2017)[,c(2:6)]
colnames(MR_1988) <- c("MM","DD","hh","ATMP1988","WTMP1988")
colnames(MR_2017) <- c("MM","DD","hh","ATMP2017","WTMP2017")

#combine 1988 and 2017 into one table
MR_1988VS2017 <- as.data.frame(left_join(MR_1988,MR_2017))
```

```{r}
#t test on air tmp
t.test(MR_1988VS2017["ATMP1988"],MR_1988VS2017["ATMP2017"])    
```
p=0 < 0.05, we can reject the null hypothesis that Air Temperature in 1988 and 2017 have the same mean

## Statistical Test- water temperature
```{r}
#t test on water tmp
t.test(MR_1988VS2017["WTMP1988"],MR_1988VS2017["WTMP2017"])    
```
p=0 < 0.05, we can reject the null hypothesis that Water Temperature in 1988 and 2017 have the same mean

## Conclusion: 
(1)ATMP at different times of a day have similar distribution
(2)WTMP at different times of a day have similar distribution
(3)Choice of the hour during a day does not make a difference
