---
title: "Analysis on Veg1.xslx"
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true 
fontsize: 8 pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(readxl)
library(stringr)
library(munsell)
library(ggplot2)
library(knitr)
library(kableExtra)


#URL of the slides to be found in the README.md file




veg.1 <- read_xlsx("veg1.xlsx")


## now get the count for each column

c <- apply(veg.1, 2, n_distinct)

e <- names(c[c>1])



veg.2 <- select(veg.1, e)



veg.3 <- dplyr::rename(veg.2, 
                       Geo = `Geo Level`, 
                       State = `State ANSI`,
                       Data = `Data Item`,
                       Category = `Domain Category`)



#let's separate Category into Label and Quant

ex1 <- separate(veg.3,Category,into = c("Label", "Quant"), sep=",")

#let's separate Quant into Type and Name

ex2 <- separate(ex1, Quant, into=c("Type","Name"), sep = ":")

ex3 <- select(ex2, -Domain)

#Let's separate and tidy up the fertilizer's names.

ex4 <- ex3 %>% separate(Label, into = c("Label","Type1"), sep=":")

ex4$Name <- with(ex4,ifelse(is.na(Type),Type1,Name))

ex4 <- select(ex4, -Type1)



#Let's focus now on tidying up the "Data" column
#The data column gives info on the kind of measurement
#that took place for a given observation

ex5 <- ex4 %>% separate(Data, into = c("name", "Measurement"), sep = "-")

ex6 <- ex5 %>% separate(Measurement, into = c("Measurement", "Info on Measurement", "Stats"), sep = ",")

#We now can discern the observation based on the measurement and its unit

#Let's rename the column Name:
#this column is equal to commodity for some observations, but not for all, so we shouldn't drop it
#as it may add some useful information
#ex7[,"Info on Commodity"] %>% unique()

ex7 <- ex6 %>% rename("Info on Commodity"=name)

ex7.1 <- ex7 %>% mutate(Geo = replace(Geo, Geo =="REGION : MULTI-STATE", "MULTI-STATE"))

#Let's split the "Name" column into Name and ID and delete the brackets

ex8 <- ex7.1 %>% separate(Name, into = c("Name", "ID"), sep = "=") %>% 
  separate(Name, into = c("Delete", "Name"), sep = "[()]") %>%
  separate(ID, into = c("ID", "Delete1"), sep = "[)]") %>%
  select(-Delete, -Delete1)

#Finally, let's replace the values in Value such as (D),(Z) and (NA) with NA

ex9 <- ex8%>% filter(!Value %in% c("(D)", "(Z)", "(NA)", NA)) %>%
  filter(!Name %in% c("TOTAL")) %>%
  filter(Stats %in% NA) %>%
  select(-Stats)

#Let's now focus our Analysis on the observation that involved the use of Restricted Chemicals

ex11 <- ex9 %>% filter(Label == "RESTRICTED USE CHEMICAL")



```



## Team members|Group 6

- Francesco Ignazio Re
- YinChia Huang 
- Yue Wu 

## Outline

- Introduction
- Data Cleaning
- Data Visualization and Exploration
- Chemicals Toxicity information
- Conclusion


## Introduction

The aim of this analysis is to give a clear understanding of the data-set veg1.xlsx. In order to do that, we intend to make the data set more intuitive to read and work on. 

###What's the data set about? {.smaller}

-The data set gathers information on the amount and type of chemicals found in food crops between the years 1990/2016 (not all the years are present), showing the location of the observations and the kind of measurement carried out. 

We would like to get a deeper insight in order to understand if the use of these chemicals inpacts significantly the quality of the crops and the land they were grown in. 




## Data Cleaning

The first thing to do is to better organize the information that we have. We make sure that every variable has a column of its own. 
Below, some examples:

```{ex1 eval=FALSE, echo= TRUE }

#Separating the column Category into Labe and Quant

ex1 <- separate(veg.3,Category,into = c("Label", "Quant"), sep=",")

#Separating the colum Quant into Type and Name

ex2 <- separate(ex1, Quant, into=c("Type","Name"), sep = ":")

#Separating chemicals name from their IDs, and deleting brackets

ex8 <- ex7.1 %>% separate(Name, into = c("Name", "ID"), sep = "=")%>% 
  separate(Name, into = c("Delete", "Name"), sep = "[()]") %>%
  separate(ID, into = c("ID", "Delete1"), sep = "[)]") %>%
  select(-Delete, -Delete1)



```

##Data Cleaning
A tricky step was understanding that many observations where just statistical results of a previous analysis. We selected them and separeted their values from the ones of the real observations.
Finally, we replace the values that lack any useful information with NAs.

```{ex1 eval=FALSE, echo= TRUE }
#We discern the observation that are statistical results.
ex6 <- ex5 %>%
  separate(Measurement,
  into = c("Measurement", "Info on Measurement", "Stats"), 
  sep = ",")

#Finally, let's replace the values such as (D),(Z) and (NA) 
#with NA and separate actual values from statistical results

ex9 <- ex8%>% filter(!Value %in% c("(D)", "(Z)", "(NA)", NA))%>%
  filter(!Name %in% c("TOTAL")) %>%
  filter(Stats %in% NA) %>%
  select(-Stats)

```


## Data Visualization and Exploration

After tidying our dataset, we are able to start visualizing the information we got, focusing on our area of interest: the amount of restricted chemicals found in the crops.

It turns out that Broccoli and Cauliflowers and the only vegetables of the data set presenting restricted chemicals. 



##Restricted Chemicals in Broccoli

```{r echo=FALSE}
#Analysis on Restricted Chemicals on Broccoli on Average

Br1 <- ex9 %>% filter(Commodity == "BROCCOLI", `Info on Measurement`==" MEASURED IN LB" , Label == "RESTRICTED USE CHEMICAL") 
Br1$Value <- as.numeric(Br1$Value)
Br1$Year <- as.numeric(Br1$Year)
Br1$Year <- as.factor(Br1$Year)

graph <- Br1 %>% group_by(Name) %>% summarize(mean(Value))

ggplot(graph, mapping=aes(x= Name, y=`mean(Value)` )) + 
  geom_bar(stat = "identity", fill="steelblue") + 
  coord_flip() +
  labs(y = "Values (LB) ", title = "Average values of restricted chemicals in broccoli ") 

```



##Restricted Chemicals in Cauliflowers

```{r echo=FALSE}
#Analysis on Restricted Chemicals on Cauliflower on Average

Cal1 <- ex9 %>% filter(Commodity == "CAULIFLOWER", `Info on Measurement`==" MEASURED IN LB" , Label == "RESTRICTED USE CHEMICAL") 

Cal1$Value <- as.numeric(Cal1$Value)
Cal1$Year <- as.numeric(Cal1$Year)
Cal1$Year <- as.factor(Cal1$Year)



graph <- Cal1 %>% group_by(Name) %>% summarize(mean(Value))


ggplot(graph, mapping=aes(x= Name, y=`mean(Value)` )) + 
  geom_bar(stat = "identity", fill="steelblue") + 
  coord_flip() +
  labs(y = "Values (LB) ", title = "Average values of restricted chemicals in calliflowers ")


```





##Values of restricted chemicals in Broccoli by year
```{r echo=FALSE}


#Analysis on Restricted Chemicals on Broccoli by year


#The graph below presents different widths according to the number of years where each chemical appears
ggplot(Br1, aes( x=Name, y= Value, fill=Year))+
  geom_bar( position = "dodge",stat="identity") + coord_flip() +  scale_fill_manual(values = c("yellow","orange", "red", "brown")) 


#the graph below works with the development version of ggplot only. It solves the difference in width of the bars

#to run it, install devtools::install_github("tidyverse/ggplot2")

#(see the script part2 for more information)


#ggplot(Br1, aes( x=Name, y= Value, fill=Year))+
 # geom_bar( position=position_dodge(preserve="single"),stat="identity") +
  #coord_flip() +  scale_fill_manual(values = c("yellow","orange", "red", "brown")) 



```


##Values of restricted chemicals in Cauliflowers by year
```{r echo=FALSE}


#plot with normal version of ggplot2

ggplot(Cal1, aes( x=Name, y= Value, fill=Year))+
 geom_bar( position = "dodge",stat="identity") + coord_flip() +  scale_fill_manual(values = c("yellow","orange", "red", "brown")) 



#the graph below works with the development version of ggplot only. It solves the difference in width of the bars

#to run it, install devtools::install_github("tidyverse/ggplot2")

#ggplot(Cal1, aes( x=Name, y= Value, fill=Year))+
 # geom_bar( position=position_dodge(preserve="single"),stat="identity") +
  #coord_flip() +  scale_fill_manual(values = c("yellow","orange", "red", "brown"))





```



##Total Average Between Broccoli and Cauliflowers 


```{r echo=FALSE}

BrCal1 <- rbind(Br1,Cal1)

grapht <- BrCal1 %>% group_by(Commodity, Name) %>% summarize(mean(Value))

ggplot(grapht, mapping=aes(x= Name, y=`mean(Value)`, fill = Commodity)) + 
 geom_bar( position="dodge",stat="identity") +
 coord_flip() +
 labs(y = "Values (LB) ", title = "Average values of restricted chemicals")


#DEV VERSION
#ggplot(grapht, mapping=aes(x= Name, y=`mean(Value)`, fill = Commodity)) + 
 # geom_bar( position=position_dodge(preserve="single"),stat="identity") +
  #coord_flip() +
  #labs(y = "Values (LB) ", title = "Average values of restricted chemicals")

```

##Chemicals toxicity information {.smaller}
Now that we have a general understanding of the quantity of each chemicals in the vegetables, we would like to know more about the chemical properties. For this reason, we gathered information on every chemical and created a table displaying:

-Soil Adsorption Coefficient: measures the amount of chemical substance adsorbed onto soil per amount of water. If the values are very low, then the soil doesn't adsorb the chemical properly and there's high risk of ground water contamination.

-LD50 on rats: value of chemical necessary to cause mortality in rats. (In the study the chemical is administered orally)

-LC50 on earthworms: concentration of chemical in air in order to cause mortality in earthworms.


##Toxicity table

Here the first rows of the table:
```{r, results='asis'}

dset1 <-read.csv("toxicity.csv")
knitr::kable(head(dset1),format = 'markdown')


```

##Analysis on toxicity by chemical in rats

```{r echo=FALSE}



df2 = read.csv("2.csv",sep = ";")
df3 = read.csv("3.csv",sep = ";")
df4 = read.csv("4.csv",sep = ";")
df5= read.csv("5.csv",sep = ";")
df6 = read.csv("6.csv",sep = ";")
df7 = read.csv("7.csv",sep = ";")
df8 = read.csv("8.csv",sep = ";")
df10 = read.csv("10.csv",sep = ";")
df11 = read.csv("11.csv",sep = ";")
df12 = read.csv("12.csv",sep = ";")
df13 = read.csv("13.csv",sep = ";")
df14 = read.csv("14.csv",sep = ";")
df15 = read.csv("15.csv",sep = ";")


df2 <- as.tibble(df2)
df3 <- as.tibble(df3)
df4 <- as.tibble(df4)
df5 <- as.tibble(df5)
df6 <- as.tibble(df6)
df7 <- as.tibble(df7)
df8 <- as.tibble(df8)
df10 <- as.tibble(df10)
df11 <- as.tibble(df11)
df12 <- as.tibble(df12)
df13 <- as.tibble(df13)
df14 <- as.tibble(df14)
df15 <- as.tibble(df15)

df2 <- df2 %>% mutate( Name= "BIFENTHRIN") %>% filter(SPECIES == "earthworm" | SPECIES == "rat" )
df3 <- df3 %>% mutate( Name= "ESFENVALERATE") %>% filter(SPECIES == "earthworm" | SPECIES == "rat")
df4 <- df4 %>% mutate( Name= "IMIDACLOPRID") %>% filter(SPECIES == "earthworm" | SPECIES == "rat")
df5 <- df5 %>% mutate( Name= "LAMBDA-CYHALOTHRIN") %>% filter(SPECIES == "earthworm" | SPECIES == "rat")
df6 <- df6 %>% mutate( Name= "METHOMYL") %>% filter(SPECIES == "earthworm" | SPECIES == "rat")
df7 <- df7 %>% mutate( Name= "NALED") %>% filter(SPECIES == "earthworm" | SPECIES == "rat")
df8 <- df8 %>% mutate( Name= "PERMETHRIN") %>% filter(SPECIES == "earthworm" | SPECIES == "rat")
df10 <- df10 %>% mutate( Name= "CHLORANTRANILIPROLE") %>% filter(SPECIES == "earthworm" | SPECIES == "rat")
df11 <- df11 %>% mutate( Name= "CHLORPYRIFOS") %>% filter(SPECIES == "earthworm" | SPECIES == "rat")
df12 <- df12%>% mutate( Name= "DIAZINON") %>% filter(SPECIES == "earthworm" | SPECIES == "rat")
df13 <- df13 %>% mutate( Name= "PRONAMIDE") %>% filter(SPECIES == "earthworm" | SPECIES == "rat")
df14 <- df14 %>% mutate( Name= "DISULFOTON") %>% filter(SPECIES == "earthworm" | SPECIES == "rat")
df15 <- df15 %>% mutate( Name= "EMAMECTIN BENZOATE") %>% filter(SPECIES == "earthworm" | SPECIES == "rat")


tox <- rbind(df2,df3,df4,df5,df6,df7,df8,df10,df11,df12,df13,df14,df15)

tox <- tox %>% filter(TYPE=="LC50" | TYPE == "LD50")


tox1 <- tox %>% filter(UNITS  == "mg/kg") %>% 
  group_by(Name,SPECIES,UNITS) %>%
  summarize(round(mean(PRIORITY)),mean(VALUES)) %>%
  rename( "Values" = `mean(VALUES)`, 
          "Priority" = `round(mean(PRIORITY))`, 
          "Units" = UNITS)

tox1 <- ungroup(tox1)

tox1$SPECIES <- as.character(tox1$SPECIES)
tox1$Units <- as.character(tox1$Units)
tox1$Values <- as.numeric(tox1$Values)



tox1 <- spread(tox1, key= SPECIES, value = Values )


tox1 <- tox1 %>% rename("Values for LD50 on rats" = rat, "Values for LC50 on earthworms"=earthworm)
tox1 %>% filter(!is.na(`Values for LD50 on rats`)) %>% ggplot(aes(x=Name, y=`Values for LD50 on rats`)) + 
  geom_bar(stat="identity",fill="steelblue") + coord_flip() +
  labs(x="Chemical",title = "Aver. Quantity of chemicals (mg/kg) causing mortality in rats")
```

##Analysis on toxicity by chemical in earthworms

```{r echo=FALSE}

tox1 %>% filter(!is.na(`Values for LC50 on earthworms`)) %>% ggplot(aes(x=Name, y=`Values for LC50 on earthworms`)) + 
  geom_bar(stat="identity",fill="steelblue") + coord_flip() +
  labs(x="Chemical",title = "Aver. Quantity of chemicals causing mortality in earthworms")
```


##Joining toxicity info to the overall table

```{r, results='asis'}

dset1 <-read.csv("finaltab.csv")
 knitr::kable(dset1, format = "markdown", table.attr = "style='width:10%;'")

```