---
title: "Analysis on Veg1.xslx"
output: 
  ioslides_presentation 
fontsize: 8 pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(readxl)
library(stringr)
library(munsell)
library(ggplot2)







veg.1 <- read_xlsx("veg1.xlsx")


## now get the count for each column

c <- apply(veg.1, 2, n_distinct)

e <- names(c[c>1])



veg.2 <- select(veg.1, e)



veg.3 <- dplyr::rename(veg.2, 
                       Geo = `Geo Level`, 
                       State = `State ANSI`,
                       Data = `Data Item`,
                       Category = `Domain Category`)



#let's separate Category into Label and Quant

ex1 <- separate(veg.3,Category,into = c("Label", "Quant"), sep=",")

#let's separate Quant into Type and Name

ex2 <- separate(ex1, Quant, into=c("Type","Name"), sep = ":")

ex3 <- select(ex2, -Domain)

#Let's separate and tidy up the fertilizer's names.

ex4 <- ex3 %>% separate(Label, into = c("Label","Type1"), sep=":")

ex4$Name <- with(ex4,ifelse(is.na(Type),Type1,Name))

ex4 <- select(ex4, -Type1)



#Let's focus now on tidying up the "Data" column
#The data column gives info on the kind of measurement
#that took place for a given observation

ex5 <- ex4 %>% separate(Data, into = c("name", "Measurement"), sep = "-")

ex6 <- ex5 %>% separate(Measurement, into = c("Measurement", "Info on Measurement", "Stats"), sep = ",")

#We now can discern the observation based on the measurement and its unit

#Let's rename the column Name:
#this column is equal to commodity for some observations, but not for all, so we shouldn't drop it
#as it may add some useful information
#ex7[,"Info on Commodity"] %>% unique()

ex7 <- ex6 %>% rename("Info on Commodity"=name)

ex7.1 <- ex7 %>% mutate(Geo = replace(Geo, Geo =="REGION : MULTI-STATE", "MULTI-STATE"))

#Let's split the "Name" column into Name and ID and delete the brackets

ex8 <- ex7.1 %>% separate(Name, into = c("Name", "ID"), sep = "=") %>% 
  separate(Name, into = c("Delete", "Name"), sep = "[()]") %>%
  separate(ID, into = c("ID", "Delete1"), sep = "[)]") %>%
  select(-Delete, -Delete1)

#Finally, let's replace the values in Value such as (D),(Z) and (NA) with NA

ex9 <- ex8%>% filter(!Value %in% c("(D)", "(Z)", "(NA)", NA)) %>%
  filter(!Name %in% c("TOTAL")) %>%
  filter(Stats %in% NA) %>%
  select(-Stats)

#Let's now focus our Analysis on the observation that involved the use of Restricted Chemicals

ex11 <- ex9 %>% filter(Label == "RESTRICTED USE CHEMICAL")



```



## Team members|Group 6

- Yue Wu 
- YinChia Huang 
- Francesco Ignazio Re


## Outline

- Introduction
- Data Cleaning
- Data Visualization
- Data Exploration
- Joining Toxicity info
- Conclusion


## Introduction

The aim of this analysis is to give a clear understanding of the data-set veg1.xlsx. In order to do that, we intend to make the data set more intuitive to read and work on. 

###What's the data set about? {.smaller}

-The data set gathers information on the amount and type of chemicals found in food crops between the years 1990/2016 (not all the years are present), showing the location of the observations and the kind of measurement carried out. 

We would like to get a deeper insight in order to understand if the use of these chemicals inpacts significantly the quality of the crops and the land they were grown in. 




## Data Cleaning

The first thing to do is to better organize the information that we have. We make sure that every variable has a column of its own. 
Below, some examples:

```{ex1 eval=FALSE, echo= TRUE }

#Separating the column Category into Labe and Quant

ex1 <- separate(veg.3,Category,into = c("Label", "Quant"), sep=",")

#Separating the colum Quant into Type and Name

ex2 <- separate(ex1, Quant, into=c("Type","Name"), sep = ":")

#Separating chemicals name from their IDs, and deleting brackets

ex8 <- ex7.1 %>% separate(Name, into = c("Name", "ID"), sep = "=")%>% 
  separate(Name, into = c("Delete", "Name"), sep = "[()]") %>%
  separate(ID, into = c("ID", "Delete1"), sep = "[)]") %>%
  select(-Delete, -Delete1)



```

##Data Cleaning
A tricky step was understanding that many observations where just statistical results of a previous analysis. We selected them and separeted their values from the ones of the real observations.
Finally, we replace the values that lack any useful information with NAs.

```{ex1 eval=FALSE, echo= TRUE }
#We discern the observation that are statistical results.
ex6 <- ex5 %>%
  separate(Measurement,
  into = c("Measurement", "Info on Measurement", "Stats"), 
  sep = ",")

#Finally, let's replace the values such as (D),(Z) and (NA) 
#with NA and separate actual values from statistical results

ex9 <- ex8%>% filter(!Value %in% c("(D)", "(Z)", "(NA)", NA))%>%
  filter(!Name %in% c("TOTAL")) %>%
  filter(Stats %in% NA) %>%
  select(-Stats)

```


## Data Visualization

After tidying our dataset, we are able to start visualizing the information we got, focusing on our area of interest: the amount of restricted chemicals found in the crops.

It turns out that Broccoli and Cauliflowers and the only vegetables of the data set presenting restricted chemicals. 



##Restricted Chemicals in Broccoli

```{r echo=FALSE}
#Analysis on Restricted Chemicals on Broccoli on Average

Br1 <- ex9 %>% filter(Commodity == "BROCCOLI", `Info on Measurement`==" MEASURED IN LB" , Label == "RESTRICTED USE CHEMICAL") 
Br1$Value <- as.numeric(Br1$Value)
Br1$Year <- as.numeric(Br1$Year)
Br1$Year <- as.factor(Br1$Year)

graph <- Br1 %>% group_by(Name) %>% summarize(mean(Value))

ggplot(graph, mapping=aes(x= Name, y=`mean(Value)` )) + 
  geom_bar(stat = "identity", fill="steelblue") + 
  coord_flip() +
  labs(y = "Values (LB) ", title = "Average values of restricted chemicals in broccoli ") 

```



##Restricted Chemicals in Cauliflowers

```{r echo=FALSE}
#Analysis on Restricted Chemicals on Cauliflower on Average

Cal1 <- ex9 %>% filter(Commodity == "CAULIFLOWER", `Info on Measurement`==" MEASURED IN LB" , Label == "RESTRICTED USE CHEMICAL") 

Cal1$Value <- as.numeric(Cal1$Value)
Cal1$Year <- as.numeric(Cal1$Year)
Cal1$Year <- as.factor(Cal1$Year)



graph <- Cal1 %>% group_by(Name) %>% summarize(mean(Value))


ggplot(graph, mapping=aes(x= Name, y=`mean(Value)` )) + 
  geom_bar(stat = "identity", fill="steelblue") + 
  coord_flip() +
  labs(y = "Values (LB) ", title = "Average values of restricted chemicals in calliflowers ")


```





##Values of restricted chemicals in Broccoli by year
```{r echo=FALSE}


#Analysis on Restricted Chemicals on Broccoli by year


#The graph below presents different widths according to the number of years where each chemical appears
ggplot(Br1, aes( x=Name, y= Value, fill=Year))+
  geom_bar( position = "dodge",stat="identity") + coord_flip() +  scale_fill_manual(values = c("yellow","orange", "red", "brown")) 


#the graph below works with the development version of ggplot only. It solves the difference in width of the bars

#to run it, install devtools::install_github("tidyverse/ggplot2")

#(see the script part2 for more information)


#ggplot(Br1, aes( x=Name, y= Value, fill=Year))+
 # geom_bar( position=position_dodge(preserve="single"),stat="identity") +
  #coord_flip() +  scale_fill_manual(values = c("yellow","orange", "red", "brown")) 



```


##Values of restricted chemicals in Caulliflowers by year
```{r echo=FALSE}


#plot with normal version of ggplot2

ggplot(Cal1, aes( x=Name, y= Value, fill=Year))+
 geom_bar( position = "dodge",stat="identity") + coord_flip() +  scale_fill_manual(values = c("yellow","orange", "red", "brown")) 



#the graph below works with the development version of ggplot only. It solves the difference in width of the bars

#to run it, install devtools::install_github("tidyverse/ggplot2")

#ggplot(Cal1, aes( x=Name, y= Value, fill=Year))+
 # geom_bar( position=position_dodge(preserve="single"),stat="identity") +
  #coord_flip() +  scale_fill_manual(values = c("yellow","orange", "red", "brown"))





```



##Total Average Between Broccoli and Calliflowers 


```{r echo=FALSE}

BrCal1 <- rbind(Br1,Cal1)

grapht <- BrCal1 %>% group_by(Commodity, Name) %>% summarize(mean(Value))

ggplot(grapht, mapping=aes(x= Name, y=`mean(Value)`, fill = Commodity)) + 
 geom_bar( position="dodge",stat="identity") +
 coord_flip() +
 labs(y = "Values (LB) ", title = "Average values of restricted chemicals")


#DEV VERSION
#ggplot(grapht, mapping=aes(x= Name, y=`mean(Value)`, fill = Commodity)) + 
 # geom_bar( position=position_dodge(preserve="single"),stat="identity") +
  #coord_flip() +
  #labs(y = "Values (LB) ", title = "Average values of restricted chemicals")

```